\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-20 doc/USER_MANUAL/11_graphics.tex
Buildbot ran the code beautifier.
updates manual using verbatim for better display in markdown

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{OUTPUT\_FILES} directory. There are several flags in the
main parameter file \texttt{Par\_file} that control the output of
these movie data files (see section \ref{cha:Main-Parameter} for
details): \texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small{} to set the
timesteps between frames, }\texttt{\small SAVE\_DISPLACEMENT}{\small{}
to save displacement instead of velocity, }\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}{\small{}
to save values at all GLL point instead of element edges. In order
to output additionally shakemaps, you would set the parameter }\texttt{\small CREATE\_SHAKEMAP}{\small{}
to }\texttt{\small .true.}{\small .}{\small \par}

@@

\begin{lyxcode}
{\footnotesize make~xcreate\_movie\_shakemap\_AVS\_DX\_GMT}{\footnotesize \par}
\end{lyxcode}

and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
in the main directory. It will create visualization files
in your format of choice. The code will prompt the user for input
parameters.

@@

When running xspecfem3D with the \texttt{\small MOVIE\_VOLUME}{\small{}
flag turned on, the code outputs several files in }\texttt{\small LOCAL\_PATH}{\small{}
specified in the main }\texttt{\small Par\_file}{\small , e.g. in
directory }\texttt{\small OUTPUT\_FILES/DATABASES\_MPI}{\small .
The output is saved by each processor at the time interval specified
by }\texttt{\small NTSTEP\_BETWEEN\_FRAMES}{\small . For all domains,
the velocity field is output to files: }\texttt{\small proc??????\_velocity\_X\_it??????.bin}{\small ,
}\texttt{\small proc??????\_velocity\_Y\_it??????.bin}{\small{} and
}\texttt{\small proc??????\_velocity\_Z\_it??????.bin}{\small . For
elastic domains, the divergence and curl taken from the velocity field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
as well: }\texttt{\small proc??????\_div\_it??????.bin}{\small , }\texttt{\small proc??????\_curl\_X\_t??????.bin}{\small ,
}\texttt{\small proc??????\_curl\_Y\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_Z\_it??????.bin}{\small .
The files denoted }\texttt{\small proc??????\_div\_}{\small ~}\\
{\small{} }\texttt{\small glob\_it??????.bin}{\small{} and }\texttt{\small proc??????\_curl\_glob\_it??????.bin}{\small{}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the }\texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small{}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).}{\small \par}

@@

To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile
it in the root directory, type:
\begin{lyxcode}
{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
\end{lyxcode}
which will create the executable \texttt{bin/xcombine\_vol\_data}.
To output the usage of this executable, type
\begin{lyxcode}
{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
\end{lyxcode}
without arguments.

@@

\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file. For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e. if you want to combine these files :

\texttt{proc000000\_velocity\_Z\_it000400.bin }

\texttt{proc000001\_velocity\_Z\_it000400.bin }

\texttt{proc000002\_velocity\_Z\_it000400.bin }

\texttt{proc000003\_velocity\_Z\_it000400.bin}

@@

\begin{lyxcode}
~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
\end{lyxcode}

Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).

@@

If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
\begin{lyxcode}
{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
\end{lyxcode}
Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
in the script according to your system.

@@

\begin{lyxcode}
{\small slice\_number.pl~Mesh\_Par\_file~output\_solver.txt~slice\_file}{\small \par}
\end{lyxcode}

which will generate a \texttt{slices\_file}.

@@

\begin{lyxcode}
{\small utils/copy\_database.pl~slice\_file~lsf\_machine\_file~filename~{[}jobid{]}}{\small \par}
\end{lyxcode}

where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
generated by the LSF scheduler, }\texttt{\small filename}{\small{}
is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
where all the kernel files are stored.}{\small \par}

@@

We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
the kernel files from all slices into one mesh file.
\begin{enumerate}
\item Compile it in the root directory:

\begin{lyxcode}
{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}

{\footnotesize ./bin/xcombine\_vol\_data~slice\_list~filename~input\_dir~output\_dir~high/low-resolution~}{\footnotesize \par}
\end{lyxcode}

where \texttt{input\_dir} is the directory where all the individual
kernel files are stored, and \texttt{output\_dir} is where the mesh
file will be written.

\item Use 1 for a high-resolution mesh, outputting all the GLL points to
the mesh file, or use 0 for low resolution, outputting only the corner
points of the elements to the mesh file.
\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
\end{enumerate}
\item \textbf{Convert mesh files into .vtu files}

@@

\begin{lyxcode}
{\footnotesize mesh2vtu.pl~-i~file.mesh~-o~file.vtu~}{\footnotesize \par}
\end{lyxcode}
\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
for its execution. Therefore, make sure you have them properly set
in the script according to your system.
\end{enumerate}
\item \textbf{Copy over the source and receiver .vtk file}

>>>>

When running \texttt{xspecfem3D} with the \texttt{MOVIE\_SURFACE}
flag turned on, the code outputs \texttt{moviedata??????} files in
the \texttt{OUTPUT\_FILES} directory. There are several flags in the
main parameter file \texttt{Par\_file} that control the output of
these movie data files (see section \ref{cha:Main-Parameter} for
details): 
\texttt{\small NTSTEP\_BETWEEN\_FRAMES} to set the
timesteps between frames, 
\texttt{\small SAVE\_DISPLACEMENT}
to save displacement instead of velocity,
\texttt{\small USE\_HIGHRES\_FOR\_MOVIES}
to save values at all GLL point instead of element edges. In order
to output additionally shakemaps, you would set the parameter
\texttt{\small CREATE\_SHAKEMAP}
to \texttt{\small .true.}.

@@

{\footnotesize
\begin{verbatim}
make xcreate_movie_shakemap_AVS_DX_GMT
\end{verbatim}
}
and run the executable \texttt{xcreate\_movie\_shakemap\_AVS\_DX\_GMT}
in the main directory. It will create visualization files
in your format of choice. The code will prompt the user for input
parameters.

@@

When running xspecfem3D with the 
\texttt{\small MOVIE\_VOLUME}
flag turned on, the code outputs several files in \texttt{\small LOCAL\_PATH}
specified in the main \texttt{\small Par\_file}, 
e.g. in directory \texttt{\small OUTPUT\_FILES/DATABASES\_MPI}.
The output is saved by each processor at the time interval specified
by 
\texttt{\small NTSTEP\_BETWEEN\_FRAMES}. 
For all domains,
the velocity field is output to files: 
{\small
\begin{verbatim}
proc??????_velocity_X_it??????.bin
proc??????_velocity_Y_it??????.bin
proc??????_velocity_Z_it??????.bin 
\end{verbatim}
}
For elastic domains, the divergence and curl taken from the velocity field,
i.e. $\nabla\cdot{\bf {v}}$ and $\nabla\times{\bf {v}}$, get stored
as well: 
{\small
\begin{verbatim}
proc??????_div_it??????.bin
proc??????_curl_X_t??????.bin
proc??????_curl_Y_it??????.bin
proc??????_curl_Z_it??????.bin
\end{verbatim}
}
The files denoted \texttt{\small proc??????\_div\_glob\_it??????.bin} and 
\texttt{\small proc??????\_curl\_glob\_it??????.bin}
are stored on the global points only, all the other arrays are stored
on all GLL points. Note that the components X/Y/Z can change to E/N/Z
according to the 
\texttt{\small SUPPRESS\_UTM\_PROJECTION}
flag (see also Appendix~\ref{cha:Coordinates} and \ref{cha:channel-codes}).

@@

To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile
it in the root directory, type:

{\footnotesize
\begin{verbatim}
make xcombine_vol_data
\end{verbatim}
}
which will create the executable \texttt{bin/xcombine\_vol\_data}.
To output the usage of this executable, type
{\footnotesize
\begin{verbatim}
./bin/xcombine_vol_data
\end{verbatim}
}
without arguments.

@@

\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file. For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e. if you want to combine these files :
{\small
\begin{verbatim}
proc000000_velocity_Z_it000400.bin
proc000001_velocity_Z_it000400.bin
proc000002_velocity_Z_it000400.bin
proc000003_velocity_Z_it000400.bin
\end{verbatim}
}

@@

{\footnotesize
\begin{verbatim}
./bin/xcombine_vol_data 0 3 velocity_Z_it000400 ./OUTPUT_FILES/DATABASES_MPI ./OUTPUT_FILES 0
\end{verbatim}
}
Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).

@@

If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView.

\noindent
For this task, you can use and modify the
script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview/}, for example:

{\footnotesize
\begin{verbatim}
mesh2vtu.pl -i velocity_Z_it000400.mesh -o velocity_Z_it000400.vtu
\end{verbatim}
}

Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
in the script according to your system.

@@

{\small
\begin{verbatim}
slice_number.pl Mesh_Par_file output_solver.txt slice_file
\end{verbatim}
}
which will generate a \texttt{slices\_file}.

@@

{\small
\begin{verbatim}
utils/copy_database.pl slice_file lsf_machine_file filename [jobid]
\end{verbatim}
}
where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
generated by the LSF scheduler, }\texttt{\small filename}{\small{}
is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
where all the kernel files are stored.}{\small \par}

@@

We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
the kernel files from all slices into one mesh file.
\begin{enumerate}
\item Compile it in the root directory:
{\footnotesize
\begin{verbatim}
make xcombine_vol_data
./bin/xcombine_vol_data slice_list filename input_dir output_dir high/low-resolution
\end{verbatim}
}
where \texttt{input\_dir} is the directory where all the individual
kernel files are stored, and \texttt{output\_dir} is where the mesh
file will be written.

\item Use 1 for a high-resolution mesh, outputting all the GLL points to
the mesh file, or use 0 for low resolution, outputting only the corner
points of the elements to the mesh file.
\item The output mesh file will have the name \texttt{filename\_rho(alpha,beta).mesh}
\end{enumerate}

\item \textbf{Convert mesh files into .vtu files}

@@

{\footnotesize
\begin{verbatim}
mesh2vtu.pl -i file.mesh -o file.vtu
\end{verbatim}
}

\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{\small utils/Visualization/Paraview/mesh2vtu} directory, which further
uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
for its execution. Therefore, make sure you have them properly set
in the script according to your system.
\end{enumerate}

\item \textbf{Copy over the source and receiver .vtk file}

### 2015-05-20 doc/USER_MANUAL/13_changing_the_model.tex
updates manual using verbatim for better display in markdown

To implement your own external tomographic model(s), you must provide
your own external tomography file(s), and choose between two possible
options:\\
 \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
= tomo}, \\
 or for more user control: \\
 \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
= default}, define the negative \texttt{material\_ID} identifier for
each element in the file \texttt{MESH/materials\_file} and use the
following format in the file \texttt{MESH/nummaterial\_velocity\_file}
when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{lyxcode}
domain\_ID~material\_ID~tomography~elastic~file\_name~1
\end{lyxcode}
where: \\
 \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\\
 \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
-1,-2,...), \\
 \indent \texttt{tomography} keyword for tomographic material definition,
\\
 \indent \texttt{elastic} keyword for elastic material definition,
\\
 \indent \texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.\\

@@

The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions. (Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
%
\begin{lyxcode}
2  -1 tomography elastic tomography\_model.xyz 1
\end{lyxcode}
%
and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}. All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.

In some cases it may be desirable to have an external tomography model that is described in more than one file. For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km. Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km. If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization. In this case  \texttt{nummaterial\_velocity\_file} is
%
\begin{lyxcode}
2  -1 tomography elastic file\_above\_15km.xyz 1 \\
2  -2 tomography elastic file\_below\_15km.xyz 1
\end{lyxcode}
%
and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.  All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth. In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{lyxcode}
call~model\_external\_values(xmesh,~ymesh,~zmesh,~~\&~~\\
~~~~~~~~~~~rho,~vp,~vs,~qmu\_atten,~iflag\_aniso,~idomain\_id)
\end{lyxcode}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

@@

To use your anisotropic models, you can either set in the \texttt{Par\_file}
the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
models may be superimposed on the mesh based upon the subroutine in
file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
The call to this subroutine is of the form
\begin{lyxcode}
call~model\_aniso(iflag\_aniso,~rho,~vp,~vs,~\&~~\\
~~~~~~~~~~~c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26,~\&~~\\
~~~~~~~~~~~c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)~
\end{lyxcode}
Input to this routine consists of:
\begin{description}
\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
generic pre-defined anisotropic models.
\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
$\rho$, $v_{p}$ and $v_{s}$.
\end{description}
Output from the routine consists of the following non-dimensional
model parameters:
\begin{description}
\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
anisotropic elastic parameters.
\end{description}
You can replace the \texttt{model\_aniso.f90} file by your own version
\textit{provided you do not change the call structure of the routine},
i.e., the new routine should take exactly the same input and produce
the required relative output.

>>>>

To implement your own external tomographic model(s), you must provide
your own external tomography file(s), and choose between two possible
options:\\
 \indent (1) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
= tomo}, \\
 or for more user control: \\
 \indent (2) set in the \texttt{Par\_file} the model parameter \texttt{MODEL
= default}, define the negative \texttt{material\_ID} identifier for
each element in the file \texttt{MESH/materials\_file} and use the
following format in the file \texttt{MESH/nummaterial\_velocity\_file}
when using CUBIT to construct your mesh (see also section \ref{subsec:Exporting-the-Mesh}):
\begin{verbatim}
domain_ID material_ID tomography elastic file_name 1
\end{verbatim}
where: \\
 \indent \texttt{domain\_ID} is 1 for acoustic or 2 for elastic materials,
\\
 \indent \texttt{material\_ID} a negative, unique identifier (i.e.,
-1,-2,...), \\
 \indent \texttt{tomography} keyword for tomographic material definition,
\\
 \indent \texttt{elastic} keyword for elastic material definition,
\\
 \indent \texttt{file\_name} the name of the tomography file and
1 a positive unique identifier.\\

@@

The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions. (Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
%
\begin{verbatim}
2  -1 tomography elastic tomography_model.xyz 1
\end{verbatim}
%
and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}. All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.

In some cases it may be desirable to have an external tomography model that is described in more than one file. For example, in cases like southern California, the length scale of variation in the structure of the wave speed model is much shorter in the sedimentary basin models within the upper 15~km. Therefore one might want to use an external tomography file that is sampled with, say, $\Delta x = 1000$~m, $\Delta y = 1000$~m, and $\Delta z = 250$~m in the uppermost 15~km, and then use $\Delta x = 2000$~m, $\Delta y = 2000$~m, and $\Delta z = 1000$~m below a depth of 15~km. If these intervals are chosen appropriately, then it will result in a pair of external tomography file that is much smaller than the alternative of having a single file with the fine discretization. In this case  \texttt{nummaterial\_velocity\_file} is
%
\begin{verbatim}
2  -1 tomography elastic file_above_15km.xyz 1
2  -2 tomography elastic file_below_15km.xyz 1
\end{verbatim}
%
and the files \texttt{file\_above\_15km.xyz} and \texttt{file\_below\_15km.xyz} will need to reside in \texttt{DATA/}.  All entries in the second column of \texttt{materials\_file} will need to be either \texttt{-1} or \texttt{-2}, depending on whether the element is above or below 15~km depth. In this sense, the construction of the mesh and the discretization of the external model will generally be done in tandem.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
                           rho, vp, vs, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

@@

To use your anisotropic models, you can either set in the \texttt{Par\_file}
the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
models may be superimposed on the mesh based upon the subroutine in
file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
The call to this subroutine is of the form
\begin{verbatim}
call model_aniso(iflag_aniso, rho, vp, vs, &
                 c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26, &
                 c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
generic pre-defined anisotropic models.
\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
$\rho$, $v_{p}$ and $v_{s}$.
\end{description}
Output from the routine consists of the following non-dimensional
model parameters:
\begin{description}
\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
anisotropic elastic parameters.
\end{description}
You can replace the \texttt{model\_aniso.f90} file by your own version
\textit{provided you do not change the call structure of the routine},
i.e., the new routine should take exactly the same input and produce
the required relative output.
